---
title: "Create a Stacked Bar Chart of Proportion"
author: "Richard White"
date: "22/04/2022"
output:
  md_document:
    variant: markdown_github
---

```{r packages, include=FALSE, message=FALSE}
for( package in c("knitr")) {
  library(package, character.only = TRUE)
}
```

```{r knitr_options, include=TRUE, message=FALSE, echo=FALSE}
opts_chunk$set(fig.width=6, fig.height=4, include=TRUE, echo=TRUE)
```

```{r load_packages}
library(ggplot2)
library(palmerpenguins)
```

Normally `geom_bar` plots the numbers of cases a each `x` position.
```{r default_bar}
ggplot(penguins, aes(x = species)) + geom_bar()
```

Adding a `fill` or `colour` aesthetic creates a stacked bar chart.
```{r stacked_bar}
ggplot(na.omit(penguins), aes(x = species, fill = sex)) + geom_bar()
```

To change it from number of cases to proportion of cases in the group, add `position = "fill"` to `geom_bar`.
```{r stacked_prop_bar}
ggplot(na.omit(penguins), aes(x = species, fill = sex)) + 
  geom_bar(position = "fill")
```

The same thing can be achieved using the `position_fill()` function. This can be useful for reversing the order of stacking, so that the legend matches the bars.
```{r flipped_stacked_prop_bar}
ggplot(na.omit(penguins), aes(x = species, fill = sex)) + 
  geom_bar(position = position_fill(reverse = TRUE)) +
  coord_flip() +
  theme(legend.position = "top")
```
